---
author: "Martin Hu"
date: ""
output: 
  html_document:
    toc: true
    toc_depth: 3
---

## Contents

This Rmd file contains test cases and expected output for the functions you are asked to write on Homework 4.  

Note that each of the test cases is randomly generated.  The `set.seed()` command at the beginning of each code block ensures that everyone's random test case is exactly the same.

## Problem 1

### Test case 1

No missing values, group variable has 2 levels.

#### Input 
```{r}
set.seed(92016)

x1 <- c(rnorm(20), 2 + rnorm(10))
groups1 <- as.factor(c(rep("A", 20), rep("B", 10)))

# Print input
x1
groups1
```

#### Expected output

```
> generateNumericSummary(x1, groups1)
$num.missing
[1] 0

$means
[1] 0.04978985 2.25163095

$sds
[1] 0.8140751 1.1815375

$p.value
[1] 0.0001297579

$var.type
[1] "numeric"
```

### Test case 2

No missing values, group variable has 4 levels.

#### Input
```{r}
set.seed(92016)

x2 <- c(rnorm(20), 2 + rnorm(10), 1 + rnorm(10), rnorm(10))
groups2 <- as.factor(c(rep("A", 20), rep("B", 10), rep("C", 10), rep("D", 10)))

# Print input
x2
groups2
```

#### Expected output
```
> generateNumericSummary(x2, groups2)
$num.missing
[1] 0

$means
[1]  0.04978985  2.25163095  0.59876555 -0.23038550

$sds
[1] 0.8140751 1.1815375 0.9309859 0.9165123

$p.value
[1] 5.355028e-07

$var.type
[1] "numeric"
```

### Test case 3

Some missing values, group variable has 2 levels.

#### Input
```{r}
x3 <- x1
x3[c(1, 5, 10, 15, 16, 22, 27, 29)] <- NA
groups3 <- groups1
  
# Print input
x3
groups3
```

#### Expected output

```
> generateNumericSummary(x3, groups3)
$num.missing
[1] 8

$means
[1] 0.1513919 2.3103344

$sds
[1] 0.8608662 1.0055923

$p.value
[1] 0.0005702583

$var.type
[1] "numeric"
```

### Test case 4

Some missing values, group variable has 4 levels.

#### Input
```{r}
x4 <- x2
x4[c(1,2,6,15,24,25,33,37,44)] <- NA
groups4 <- groups2

# Print input
x4
groups4
```

#### Expected output

```
> generateNumericSummary(x4, groups4)
$num.missing
[1] 9

$means
[1]  0.03078715  2.38988151  0.70921024 -0.27178988

$sds
[1] 0.7536143 1.2947970 0.9746499 0.9621372

$p.value
[1] 4.077215e-06

$var.type
[1] "numeric"
```

## Problem 2

### Test case 1

No missing values, group variable has 2 levels.

#### Input 
```{r}
set.seed(92016)

b1 <- c(round(runif(20)), as.numeric(runif(10) < 0.8))
groups1 <- as.factor(c(rep("A", 20), rep("B", 10)))

# Print input
b1
groups1
```

#### Expected output

```
> generateBinarySummary(b1, groups1)
$num.missing
[1] 0

$prop
[1] 0.6 0.8

$p.value
[1] 0.4194711

$var.type
[1] "binary"
```

### Test case 2

No missing values, group variable has 4 levels.

#### Input
```{r}
set.seed(92016)

b2 <- c(round(runif(20)), 
        as.numeric(runif(10) < 0.8), 
        as.numeric(runif(10) < 0.2),
        as.numeric(runif(10) < 0.4))
groups2 <- as.factor(c(rep("A", 20), rep("B", 10), rep("C", 10), rep("D", 10)))

b2
groups2
```

#### Expected output
```
> generateBinarySummary(b2, groups2)
$num.missing
[1] 0

$prop
[1] 0.6 0.8 0.0 0.3

$p.value
[1] 0.0006143226

$var.type
[1] "binary"
```

### Test case 3

Some missing values, group variable has 2 levels.

#### Input
```{r}
b3 <- b1
b3[c(1, 5, 10, 15, 16, 22, 27, 29)] <- NA
groups3 <- groups1

# Print input
b3
groups3
```

#### Expected output

```
> generateBinarySummary(b3, groups3)
$num.missing
[1] 8

$prop
[1] 0.5333333 0.7142857

$p.value
[1] 0.6478328

$var.type
[1] "binary"
```

### Test case 4

Some missing values, group variable has 4 levels.

#### Input
```{r}
b4 <- b2
b4[c(1,2,6,15,24,25,33,37,44)] <- NA
groups4 <- groups2

# Print input
b4
groups4
```

#### Expected output

```
> generateBinarySummary(b4, groups4)
$num.missing
[1] 9

$prop
[1] 0.6875000 0.7500000 0.0000000 0.3333333

$p.value
[1] 0.002827139

$var.type
[1] "binary"
```

## Problem 3 and 3b

Note that the `generateVariableSummary` function should product the exact same output as the functions in Problem 1 and Problem 2 on the test cases provided for those functions.  You should therefore test your generateVariableSummary function on all 8 test cases given above. 

Here are a couple of test cases that are not covered by the 8 you already have.  Answers are provided both for the function in Problem 3 and the improved function in Problem 3b.

### Test case 1


#### Input
```{r}
set.seed(92016)
y1 <- as.factor(sample(c("cat", "dog", "bird"), 30, replace = TRUE))

# Print input
y1
```

#### Expected output: Problem 3
```
generateVariableSummary(y1, groups1)
NULL
```

#### Expected output: Problem 3b
```
> generateVariableSummary(y1, groups1)
$xbird
$xbird$num.missing
[1] 0

$xbird$prop
[1] 0.45 0.30

$xbird$p.value
[1] 0.6941529

$xbird$var.type
[1] "binary"


$xcat
$xcat$num.missing
[1] 0

$xcat$prop
[1] 0.1 0.6

$xcat$p.value
[1] 0.007234844

$xcat$var.type
[1] "binary"


$xdog
$xdog$num.missing
[1] 0

$xdog$prop
[1] 0.45 0.10

$xdog$p.value
[1] 0.1007607

$xdog$var.type
[1] "binary"


$var.type
[1] "factor"
```

### Test case 2

#### Input
```{r}
set.seed(92016)
y2 <- sample(c(1, 12, 0, "asdf"), 50, replace = TRUE)

# Print input
y2
```

#### Expected output (same for 3 and 3b)

```
> generateVariableSummary(y2, groups2)
NULL
```

## Problem 4

The test cases defined above are sufficient for testing this function.  I therefore display only the expected outputs.

### Test case 1: x1, groups1

```
> formatVariableSummary(generateVariableSummary(x1, groups1))
[1] "0"           "0.05 (0.81)" "2.25 (1.18)" "1e-04"  
```

### Test case 2: b1, groups1

```
> formatVariableSummary(generateVariableSummary(b1, groups1))
[1] "0"      "60%"    "80%"    "0.4195"

```
### Test case 3: x2, groups2

```
> formatVariableSummary(generateVariableSummary(x2, groups2))
[1] "0"            "0.05 (0.81)"  "2.25 (1.18)"  "0.6 (0.93)"  "-0.23 (0.92)" "0"     

```

### Test case 4: b2, groups2

```
> formatVariableSummary(generateVariableSummary(b2, groups2))
[1] "0"     "60%"   "80%"   "0%"    "30%"   "6e-04"

```

### Test case 5: x3, groups1

```
> formatVariableSummary(generateVariableSummary(x3, groups1))
[1] "8"           "0.15 (0.86)" "2.31 (1.01)" "6e-04"  

```

### Test case 6: b3, groups1

```
> formatVariableSummary(generateVariableSummary(b3, groups1))
[1] "8"      "53.3%"  "71.4%"  "0.6478"

```

### Test case 7: x4, groups2

```
> formatVariableSummary(generateVariableSummary(x4, groups2))
[1] "9"            "0.03 (0.75)"  "2.39 (1.29)"  "0.71 (0.97)"  "-0.27 (0.96)" "0"  

```

### Test case 8: b4, groups2

```
> formatVariableSummary(generateVariableSummary(b4, groups2))
[1] "9"      "68.8%"  "75%"    "0%"     "33.3%"  "0.0028"

```

### Test case 9: y1, groups1

```
> formatVariableSummary(generateVariableSummary(y1, groups1))
      [,1] [,2]  [,3]  [,4]    
xbird "0"  "45%" "30%" "0.6942"
xcat  "0"  "10%" "60%" "0.0072"
xdog  "0"  "45%" "10%" "0.1008"
```

### Test case 10: y2, groups2

```
> formatVariableSummary(generateVariableSummary(y2, groups2))
NULL
```

## Problem 5

### Test case 1

Grouping variable has two levels.

#### Input
```{r}
# form test cases into a data frame
df1 <- data.frame(x1, x3, b1, b3, y1, groups1)

var.names <- c("x1", "x3", "y1", "b1", "b3")
group.name <- "groups1"

# Print first bit of input
head(df1)
```

#### Expected output

```
> generateDataSummary(df1, var.names, group.name)
     Variable  Missing A             B             P-value 
[1,] "x1"      "0"     "0.05 (0.81)" "2.25 (1.18)" "1e-04" 
[2,] "x3"      "8"     "0.15 (0.86)" "2.31 (1.01)" "6e-04" 
[3,] "y1:bird" "0"     "45%"         "30%"         "0.6942"
[4,] "y1:cat"  "0"     "10%"         "60%"         "0.0072"
[5,] "y1:dog"  "0"     "45%"         "10%"         "0.1008"
[6,] "b1"      "0"     "60%"         "80%"         "0.4195"
[7,] "b3"      "8"     "53.3%"       "71.4%"       "0.6478"
```

### Test case 2

Grouping variable has 4 levels

#### Input
```{r}
# form test cases into a data frame
df2 <- data.frame(x2, x4, b2, b4, y2, groups2)

var.names2 <- c("x2", "x4", "y2", "b2", "b4")
group.name2 <- "groups2"

# Print first bit of input
head(df2)
```

#### Expected output

```
> generateDataSummary(df2, var.names2, group.name2)
     Variable  Missing A             B             C             D              P-value 
[1,] "x2"      "0"     "0.05 (0.81)" "2.25 (1.18)" "0.6 (0.93)"  "-0.23 (0.92)" "0"     
[2,] "x4"      "9"     "0.03 (0.75)" "2.39 (1.29)" "0.71 (0.97)" "-0.27 (0.96)" "0"     
[3,] "y2:0"    "0"     "20%"         "10%"         "10%"         "30%"          "0.6101"
[4,] "y2:1"    "0"     "10%"         "40%"         "0%"          "20%"          "0.0696"
[5,] "y2:12"   "0"     "30%"         "20%"         "50%"         "30%"          "0.5807"
[6,] "y2:asdf" "0"     "40%"         "30%"         "40%"         "20%"          "0.7825"
[7,] "b2"      "0"     "60%"         "80%"         "0%"          "30%"          "6e-04" 
[8,] "b4"      "9"     "68.8%"       "75%"         "0%"          "33.3%"        "0.0028"
```
