---
author: "Martin (Yuxiang) Hu"
output: 
  html_document:
    theme: paper
    highlight: tango
    toc: true
    toc_depth: 3
    fig_width: 5
    fig_height: 5
---

### Homework outline

This homework is designed to give you practice with writing a more complex function.  In particular, your final function will make calls to other simpler functions that you write along the way.  The overall task has been broken up into sub-tasks to help you decompose the problem into smaller, more tractable components.

> **Goal**: You will produce a function that generates what's often called 'Table 1'.  This table is intended to provide summary information about the variables in your data, broken down by some factor of interest.  

Problem 5 will have you apply the function to the `titanic3` data, which is available in the `PASWR` library.

```{r}
library(PASWR)
str(titanic3)
library(plyr)
```

You'll use your function on the `titanic3` data to produce the following table, which summarizes information about passengers on the Titanic, displaying summaries separately depending on the passenger's fare class (1st, 2nd or 3rd).

|Variable                  |Missing |1st           |2nd           |3rd           |P-value |
|:-------------------------|:-------|:-------------|:-------------|:-------------|:-------|
|Survival rate             |0       |61.9%         |43%           |25.5%         |0       |
|Sex:female                |0       |44.6%         |38.3%         |30.5%         |0       |
|Sex:male                  |0       |55.4%         |61.7%         |69.5%         |0       |
|Age                       |263     |39.16 (14.55) |29.51 (13.64) |24.82 (11.96) |0       |
|# siblings/spouses aboard |0       |0.44 (0.61)   |0.39 (0.59)   |0.57 (1.3)    |0.0279  |
|# children/parents aboard |0       |0.37 (0.72)   |0.37 (0.69)   |0.4 (0.98)    |0.779   |
|Fare ($)                  |1       |87.51 (80.45) |21.18 (13.61) |13.3 (11.49)  |0       |
|Embarked:                 |0       |0.6%          |0%            |0%            |0.1054  |
|Embarked:Cherbourg        |0       |43.7%         |10.1%         |14.2%         |0       |
|Embarked:Queenstown       |0       |0.9%          |2.5%          |15.9%         |0       |
|Embarked:Southampton      |0       |54.8%         |87.4%         |69.8%         |0       |

From this table we can conclude things such as:

  - The Fare was unknown for just 1 of the passengers
  - There is a statistically significant difference in the proportion of female passengers across the fare classes (p-value < 0.0001).  1st class had 44.6% female passengers, compared to 30.5% in 3rd class.

### Important notes

#### Handling missing values

You will be encountering missing values in this homework (entries coded `NA`).  To make sure that not all of your calculations turn out to be `NA`, you'll need to be careful about how you treat missing values.

Example:  When using the `mean()` function, you'll want to use `mean(..., na.rm=TRUE)`, or some alternative.  

```{r}
x <- c(NA, 1, 5, 3, NA, 2)
mean(x)
mean(x, na.rm=TRUE) # Remove missing values
mean(x[!is.na(x)]) # Select only non-missing values
```

Functions like `t.test()` have an `na.action` argument.  The default is to use whatever is returned by `getOption("na.action")`.  You'll want this to be `"na.omit"`, which is the default.

#### Testing your functions

Just as with Homework 3, I will be providing you with a list of test cases on which to test your function.  The test cases and printouts of expected output are available in a supplementary document on the course website.

### Problem 1: Summarizing numeric data

Write a function with the following specification:

1. Input: 2 inputs
    - `x`: numeric vector
    - `groups`: factor vector of the same length as `x`

2. Output: a list with the following elements:
    - `num.missing`: the number of values of `x` that are missing
    - `means`: means of `x` across the different levels of `groups`
    - `sds`: standard deviations of `x` across the different levels of `groups`
    - `p.value`: p-value output by a test of whether the mean of `x` is the same across all levels of `groups`  (i.e., p-value from a t-test or 1-way ANOVA)
    - `var.type = "numeric"`: a field called `var.type`, which is set to the string `"numeric"`
    
Here is an example of the output (in this example, the input had 2 missing values, and `groups` had 3 levels)

```
$num.missing
[1] 2

$means
[1]  0.06646450 -0.01969902  1.96278957

$sds
[1] 0.5800607 1.4293022 0.7123807

$p.value
[1] 0.01107936

$var.type
[1] "numeric"
```

```{r}
generateNumericSummary <- function(x, groups) {
  # Edit me
  my.list <- list(num.missing = sum(is.na(x)), 
                  means = aggregate(x, by = list(groups), FUN = mean, na.rm = TRUE)[,2],
                  sds = aggregate(x, by = list(groups), FUN = sd, na.rm = TRUE)[,2],
                  p.value = summary(aov(x ~ groups),na.action = na.omit)[[1]][["Pr(>F)"]][[1]],
                  var.type = "numeric"
                  )
  return(my.list)
}
```

**Hint:** Missing values in R are coded as `NA`.  You can use the function `is.na(x)` to test if `x` is missing.

### Problem 2: Summarizing binary data

Write a function with the following specification:

1. Input: 2 inputs
    - `x`: binary (0/1) vector
    - `groups`: factor vector of the same length as `x`

2. Output: a list with the following elements:
    - `num.missing`: the number of values of `x` that are missing
    - `prop`: the proportion of non-NA `x` values that equal 1 in each level of `groups`
    - `p.value`: p-value output by a test of whether the proportion of `x` that is equal to 1 is the same across all levels of `groups`  (i.e., p-value from a fisher.test)
    - `var.type = "binary"`: a field called `var.type`, which is set to `"binary"`

Here is an example of the output (in this example, the input had 4 missing values, and `groups` had 3 levels)
```
$num.missing
[1] 4

$prop
[1] 0.3333333 0.0000000 0.6000000

$p.value
[1] 0.2525253

$var.type
[1] "binary"
```

```{r}
generateBinarySummary <- function(x, groups) {
  # Edit me
  my.list <- list(num.missing = sum(is.na(x)), 
                  prop = aggregate(x, by = list(groups), 
                                   FUN = function(x){
                                     y = sum(!is.na(x) & x == 1)/ sum(!is.na(x));
                                     return(y)
                                   })[,2],
                  p.value = fisher.test(table(x,groups))$p.value,
                  var.type = "binary"
                  )
  return(my.list)
}
```

### Problem 3: Getting the right data type

Write a function with the following specification:

1. Input: 2 inputs
    - `x`: a vector
    - `groups`: factor vector of same length as `x`

2. Output:
    - If `x` is binary, return `generateBinarySummary(x, groups)`
    - Else, if `x` is numeric, return `generateNumericSummary(x, groups)`
    - Else, return `NULL`
    
Here's an example of output when `x` is numeric and `groups` has 2 levels
```
$num.missing
[1] 2

$means
[1] 0.02816961 1.96278957

$sds
[1] 0.9676708 0.7123807

$p.value
[1] 0.001395812

$var.type
[1] "numeric"
```

Here's an example of output when `x` is binary and `groups` has 2 levels

```
$num.missing
[1] 4

$prop
[1] 0.1428571 0.6000000

$p.value
[1] 0.2222222

$var.type
[1] "binary"
```
    
```{r}
# Use the function created in homework3.
isBinary <- function(x, allow.NA = FALSE) {
  if (allow.NA) {
    if (sum(is.na(x)) == 0) {
      if (sum(x == 0) + sum (x == 1) == length(x)) {
        return(TRUE)
      } else {
        return(FALSE)
      }
    } else {
      if (sum(x == 0 & !is.na(x)) + sum (x == 1 & !is.na(x)) + sum(is.na(x)) == length(x)) {
        return(TRUE)
      } else {
        return(FALSE)
      }
    }
  } else {
    if (sum(is.na(x)) == 0) {
      if (sum(x == 0) + sum (x == 1) == length(x)) {
        return(TRUE)
      } else {
        return(FALSE)
      }
    } else {
      return(FALSE)
    }
  }
}

generateVariableSummary <- function(x, groups) {
  # Edit me
  if (isBinary(x, TRUE)) {
    generateBinarySummary(x, groups)
  } else if (is.numeric(x)) {
    generateNumericSummary(x, groups)
  } else {
    return(NULL)
  }
}
```

### Problem 3b: Dealing with factors that have more than 2 levels

This problem is a continuation of Problem 3.  In this part of the problem you will improve your `generateVariableSummary` function to handle factors with 2 or more levels.  I provide you with some starter code that takes a k-level factor and converts it to a matrix with k columns of binary vectors.  This is called the **binary representation of `x`**.   Once you have a factor represented in this way, you are able to apply the `generateBinarySummary` function to each column to produce a summary of the factor variable.

```{r}
# Example code
# Start with a factor.  This one has 3 levels.
x.factor <- c("cat", "dog", "bird", "bird", "dog", "cat", "dog")
# Binary representation of x
x.matrix <- model.matrix( ~ x.factor - 1)
x.matrix
```

Modify your function from Problem 3 to have the following specification:

1. Input: 2 inputs
    - `x`: a vector
    - `groups`: factor vector of same length as `x`

2. Output:
    - If `x` is a k-level factor, return a list with k+1 elements. 
        - The first k elements should be the results of running `generateBinarySummary` on every column of the binary representation of `x`.
        - The last element should be called `var.type` and should have the value `"factor"`.
    - Else, if `x` is binary, return `generateBinarySummary(x, groups)`
    - Else, if `x` is numeric, return `generateNumericSummary(x, groups)`
    - Else, return `NULL`
    
Here is an example of the desired output (in this example, the input had 1 missing value, was a factor with 3 levels, and `groups` had 2 levels)

```
> x
 [1] cat  bird cat  bird bird dog  dog  cat  bird dog  bird cat  cat  cat 
[15] cat  bird bird dog  cat  cat 
Levels: bird cat dog
> groups
 [1] 1 1 1 0 1 1 1 1 0 0 1 0 1 0 1 1 1 1 0 1
Levels: 0 1

# Output:
$xbird
$xbird$num.missing
[1] 0

$xbird$prop
[1] 0.3333333 0.3571429

$xbird$p.value
[1] 1

$xbird$var.type
[1] "binary"


$xcat
$xcat$num.missing
[1] 0

$xcat$prop
[1] 0.5000000 0.4285714

$xcat$p.value
[1] 1

$xcat$var.type
[1] "binary"


$xdog
$xdog$num.missing
[1] 0

$xdog$prop
[1] 0.1666667 0.2142857

$xdog$p.value
[1] 1

$xdog$var.type
[1] "binary"


$var.type
[1] "factor"
```

```{r}
generateVariableSummary <- function(x, groups) {
  # Edit me
  final.list <- list()
  if (is.factor(x)) {
    x.matrix <- model.matrix( ~ x - 1)
    for (i in 1:ncol(x.matrix)){
      if (isBinary(c(x.matrix[,i]), TRUE)) {
        result.list <- generateBinarySummary(c(x.matrix[,i]), groups)
      } else {
        result.list <- NULL
      }
      final.list[[i]] <- result.list
      names(final.list)[i] <- colnames(x.matrix)[i]
    }
    final.list[["var.type"]] <- "factor"
  } else if (isBinary(x, TRUE)){
    final.list <- generateBinarySummary(x, groups)
  } else if (is.numeric(x)){
    final.list <- generateNumericSummary(x, groups)
  } else {
    final.list <- NULL
  }
  return(final.list)
}
```

### Problem 4: Formatting output

Write a function with the following specification:

1. Input: 1 input
   - `var.summary`: the output of a `generateVariableSummary` call

2. Output: 
   - If `var.summary` is for a numeric variable and the grouping variable has `k` levels, return a vector of the form
   (`num.missing`, `"mean[1] (sd[1])"`, `"mean[2] (sd[2])"`, ..., `"mean[k] (sd[k])"`, `p.value`)
   
   - If `var.summary` is for a binary variable and the grouping variable has `k` levels, return a vector of the form
   (`num.missing`, `"prop[1]%"`, `"prop[2]%"`, ..., `"prop[k]%"`, `p.value`)
   
   - If `var.summary` is for a factor variable with `j` levels and the grouping variable has `k` levels, you should return a `j x k` table where the `j`th row of the table is the vector (`num.missing`, `"prop[1]%"`, `"prop[2]%"`, ..., `"prop[k]%"`, `p.value`) corresponding to the `j`th level of the factor.  
   
Here's sample output when `x` is numeric and `groups` has 3 levels

```
[1] "0"           "0.23 (0.91)" "1.3 (0.81)"  "0.93 (1.2)"  "0.2218"  
```

Here's sample output when `x` is binary and `groups` has 2 levels

```
"4"      "14.3%"  "60%"    "0.2222"
```

Here's sample output wen `x` is a 3-level factor and `groups` has 2 levels
```
     [,1] [,2]    [,3]    [,4]
[1,] "0"  "33.3%" "35.7%" "1" 
[2,] "0"  "50%"   "42.9%" "1" 
[3,] "0"  "16.7%" "21.4%" "1" 
```

```{r}
formatVariableSummary <- function(var.summary) {
  # Edit me
  my.vector <- c()
  if (is.null(var.summary)) {
    my.vector <- NULL
  } else if (var.summary[["var.type"]] == "numeric") {  # This section is for numeric
    my.vector <- c(my.vector, var.summary[["num.missing"]])
    for (i in 1:length(var.summary[["means"]])) {
      my.vector <- c(my.vector, paste(round(var.summary[["means"]][[i]],2),
                                      " (", 
                                      round(var.summary[["sds"]][[i]],2),
                                      ")",
                                     sep = ""))
    }
    my.vector <- c(my.vector, format(round(var.summary[["p.value"]],4),scientific = FALSE))
  } else if (var.summary[["var.type"]] == "binary") {  # This section is for binary
    my.vector <- c(my.vector, var.summary[["num.missing"]])
    for (i in 1:length(var.summary[["prop"]])) {
      my.vector <- c(my.vector, paste(round(var.summary[["prop"]][[i]]*100,1), "%", sep = ""))
    }
    my.vector <- c(my.vector, format(round(var.summary[["p.value"]],4),scientific = FALSE))
  } else if (var.summary[["var.type"]] == "factor") {  # This section is for factor
    for (j in 1:(length(var.summary) - 1)) {
      row.vector <- NULL
      row.vector <- c(row.vector, var.summary[[j]][["num.missing"]])
      for (i in 1:length(var.summary[[j]][["prop"]])) {
        row.vector <- c(row.vector, paste(round(var.summary[[j]][["prop"]][[i]]*100,1), "%", sep = ""))
      }
      row.vector <- c(row.vector, format(round(var.summary[[j]][["p.value"]],4),scientific = FALSE))
      my.vector <- rbind(my.vector, row.vector)
      rownames(my.vector)[[j]] <- names(var.summary)[[j]]
    }
  } else {
    my.vector <- NULL
  }
  return(my.vector)
}
```

**Hints:**  Given two numbers `a` and `b`, here's how you get the formatting for `"mean[1] (sd[1])"`.  This gives an example where the mean is 21.89 and the sd is 1.83.

```{r}
a <- 21.89
b <- 1.83
paste(a, " (", b , ")", sep="")
```

Similarly, given a percentage `a`, here's how to get R to return `"a%"`:

```{r}
a <- 75.2
paste(a, "%", sep="")
```

Alternatively, if you're familiar with `sprintf()` from C or other languages, you are free to use that.

### Problem 5: Piecing things together

**(a)** Using the functions you write in Problems 1 - 4, write a function with the following specification:

1. Input: 3 inputs
    - `dataset`: a data frame
    - `var.names`: a character vector that is a subset of `colnames(dataset)`
    - `group.name`: a string that gives the name of the grouping variable
   
2. Output:
    - Your function should return a data frame that has the following form (this is an example where we have three variables, one numeric and two binary, and a **group variable with 2 levels**):
    
|  Variable    | Missing | groups[1] | groups[2]  | P-value |
|:------------:|:-------:|:---------:|:----------:|:-------:|
| numeric.var  |   12    |45.2 (2.7) | 53.9 (4.1) |  0.0142 |
| binary.var1  |    0    |  14%      |   21%      |  0.0679 |
| binary.var2  |    7    |  95%      |   94%      |  0.7810  |

  - The data frame will have 3 + #(levels in grouping variable) columns
  - The first column will give the variable names specified in `var.names`
  - The second column will give the number of missing values
  - The next #(levels in grouping variable) columns will give the group means and standard deviations (as shown) if the variable is **numeric**, or % of 1's if the variable is **binary**.
  - The last column will give the p-value for testing whether the group means (or proportions) are the same across all levels of the grouping variable.
  
  
```{r}
generateDataSummary <- function(dataset, varnames, group.name) {
  # Edit me
  my.vector <- NULL
  for (i in 1:length(varnames)) {
    row.vector <- NULL
    row.vector <- formatVariableSummary(generateVariableSummary(dataset[[varnames[i]]],
                                                                dataset[[group.name]]))
    if (is.matrix(row.vector)) {
      my.vector <- rbind(my.vector, cbind(Variables = paste(varnames[i],
                                                            substring(rownames(row.vector),2),
                                                            sep = ":"), 
                                          row.vector))
    } else {
      my.vector <- rbind(my.vector, c(Variables = varnames[i], row.vector))
    }
  }
  colnames(my.vector) <- c("Variable","Missing",levels(dataset[[group.name]]),"P-value")
  rownames(my.vector) <- NULL
  return(my.vector)
}
```

**Hint**: One way of doing this is to use a loop and the `rbind()` function to build up your summary table row by row.

**(b)** Using the `kable` function, create a nice summary table obtained by running `generateDataSummary()` on the `titanic3` data from the `PASWR` library.  I've gotten you started by defining some variables.  


```{r}
# Load libraries
library(PASWR)
library(plyr)
library(knitr)

# Use the following var names to form your table:
var.names <- c("survived", "sex", "age", "sibsp", "parch", "fare", "embarked")

# Use passenger glass (pclass) as your grouping variable.
group.name <- "pclass"
```

```{r, results = 'asis'}
  # Edit me
kable(generateDataSummary(titanic3,var.names,group.name), format = 'markdown')
```


**(c)** Look at the survival rate row of your table and answer the following questions.

- What passenger class has the highest survival rate?  
     - *1st*

- What is the survival rate for this passenger class?  
     - *61.9%*
     
- Are survival rates statistically significantly different across the passenger classes?  
     - *According to the p-value, yes.*

- What statistical test was used to obtain the p-value?
     - *It adopts fisher test.*

**(d)** Look at the age row of your table and answer the following questions.  

- For how many passengers in the data is the age unknown?  
     - *263*

- Is there a statistically significant difference in average passenger age across the passenger classes?  
     - *According to the p-value, yes.*

- What statistical test was used to obtain the p-value?
     - *It adopts one way ANOVA test*
